---
title: "Intro to 'testthat' package"
author: "Niels Aka"
date: "27 September 2016"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
suppressPackageStartupMessages(library(testthat))
```

## Why (unit) test?

[Wikipedia](https://en.wikipedia.org/wiki/Unit_testing): "unit testing is a software testing method by which individual units of source code [...] are tested to determine whether they are fit for use" \newline
$\Rightarrow$ lowest level of testing
\newline

- ensures that single functions work as expected

- catch bugs early on

- save time later on

## Why test as a scientist?

- a plus/minus sign error can invalidate research results...

- more integrity, reliability, trust

## Why test in R?

- starts to make sense when you
    
    * write own functions
    
    * customise other functions
  
- R language is flexible (implicit coercion, recycling) \newline
$\Rightarrow$ make sure inputs + output are as expected

## Unit testing in R: the `testthat` package

- ideal for packages, but works also with scripts

- tests are organised in three levels
    1. context  $\rightarrow$ general functionality
    2. test suites $\rightarrow$ a specific aspect
    3. expectations $\rightarrow$ a single unit of functionality

## Simple example
```{r, eval=FALSE}
context("String length")

test_that("str_length is number of characters", {
  expect_equals(str_length("a"), 1)
  expect_equals(str_length("ab"), 2)
  expect_equals(str_length("abc"), 3)
})
```

## Workflow

When using `testthat`, there are usually three steps

1. write code
2. write tests
3. source `testthat.R` which runs all tests in separate testing environments
  and reports summary of successes and failures

## Example file: `transform.R`
```{r}
# transform time-series in levels into growth rates
trans_gr <- function(x) {
  diff(log(x))
}
```

## Example file: `test-transform.R`
```{r, eval=FALSE}
context("Testing transformation routines")

test_that("transforming to growth rates works", {
  n    <- 5
  rate <- 0.02
  x <- (1 + rate) ^ (1:n)
  
  # test class
  expect_is(object = trans_gr(x),
            class = "numeric")
  # test numerical results
  ref <- rep(rate, n - 1)
  expect_equal(object = trans_gr(x), expected = ref,
               tol = 1e-3)
  expect_identical(object = trans_gr(x), expected = ref)
})
```

## Example file: `testthat.R`

```{r, echo=TRUE}
source("transform.R")
testthat::test_file("test-transform.R")
```


## Resources

- [R Journal entry](https://www.google.de/url?sa=t&rct=j&q=&esrc=s&source=web&cd=1&cad=rja&uact=8&ved=0ahUKEwj8hJqe0a_PAhVpBMAKHWr4D10QFggjMAA&url=https%3A%2F%2Fjournal.r-project.org%2Farchive%2F2011-1%2FRJournal_2011-1_Wickham.pdf&usg=AFQjCNGMHc4ywSRYqlG_4qqtvXNf1mxFCg&bvm=bv.133700528,d.ZGg)

- [CRAN website](https://cran.r-project.org/web/packages/testthat/index.html)

- [Hadley's website](http://r-pkgs.had.co.nz/tests.html)
