---
title: "Escaping in R"
author: "Niels Aka"
date: "10 November 2017"
output: html_document
---

```{r, echo=FALSE, include=FALSE}
Sys.setenv(LANG = "en")
```

When and why to use how many backslashes in R? That is not always
straightforward to say. A backslash inside a string has a special meaning
for R. If it sees one, this is a sign to "start an escape sequence inside character constants" and attach special meaning to the character following the backslash. The quote is taken from `?Quotes`. The help page provides a list with possible escape sequences such as `\n` for newline or `\\` for
using an actual backslash.

Here is an example.

```{r}
cat("zz \n zz")
```

Adding another backslash will escaping the second backslash followed by a normal `"n"`. So then,

```{r}
cat("zz \\n zz")
```

If we add another backslash, the first two will result in a single backslash in our string, now followed by the special symbol for newline.

```{r}
cat("zz \\\n zz")
```



`?Quotes` also says

> Escaping a character not in the following table is an error.

So,

```{r, error = TRUE}
cat("zz \A zz")
```
```{r}
cat("zz \\A zz")
```
```{r, error = TRUE}
cat("zz \\\A zz")
```

Regular expressions
===================

So far so good. Things start to get a bit more complicated when you work with
regular expressions and what to, for example, substitute parts of a string
with another string.

```{r}
gsub(pattern = "Y", replacement = "X", x = "zz Y zz")
```

Let's wrap that line of code in a function to concentrate only on the string
we want to replace `"Y"`with.

```{r}
my_sub <- function(rep) {
  cat(gsub(pattern = "Y", replacement = rep, x = "zz Y zz"))
}
```

Same example as before.

```{r}
my_sub("\n")
```

Looks good. But the behaviour starts to be different when adding more backslashes.

```{r}
my_sub("\\n")
```

Instead of `zz \n zz` we now have `zz n zz`. The backslash that we inserted in the string by escaping it did not get printed.

```{r}
my_sub("\\")
```

That means we have to add more backslashes than before to get the desired result.

```{r}
my_sub("\\\n")

my_sub("\\\\n")
```

Hence, one [R wisdom](https://stackoverflow.com/questions/14879204/how-to-escape-a-backslash-in-r) on stack overflow 

    When in doubt, keep adding slashes until it works. – joran Feb 14 '13 at 16:21 

But to understand better what is going one, let's assume we wanted to replace every backslash in a string with two more backslashes.

```{r, error = TRUE}
gsub("\", "\\", "\")
```

Ah, yes, forgot to escape the backslashes. They were escaping the quotation marks instead.

```{r, error = TRUE}
# two backslashes means string of length one
# containing one backslash.
string <- "\\"
# we are looking for a pattern consisting of
# one backslash
pattern <- "\\"
# and want to replace with two
replacement <- "\\\\"

gsub(pattern, replacement, string)
```

Something is still off. Apparently the pattern we are looking for is not a valid regular expression.

```{r}
string <- "\\"
pattern <- "\\\\"
replacement <- "\\\\"

(x <- gsub(pattern, replacement, string))
nchar(x)
```

Hm, the code ran. But the outcome is still of length one.

```{r}
string <- "\\"
pattern <- "\\\\"
replacement <- "\\\\\\\\" # add more backslashes in replacement

(x <- gsub(pattern, replacement, string))
nchar(x)
```
Now it finally worked! This example was taken from [here](https://stackoverflow.com/questions/27491986/r-gsub-replacing-backslashes) with the following insight:

    Here's what you need:
    
    gsub("\\\\", "\\\\\\\\", "\\")
    [1] "\\\\"
    
    The reason that you need four backslashes to represent one literal
    backslash is that "\" is an escape character in both R strings and for the
    regex engine to which you're ultimately passing your patterns. If you were
    talking directly to the regex engine, you'd use "\\" to indicate a literal
    backslash. But in order to get R to pass "\\" on to the regex engine, you
    need to type "\\\\". 

## Background Material & Links

* Short [Tutorial](https://rstudio-pubs-static.s3.amazonaws.com/74603_76cd14d5983f47408fdf0b323550b846.html) on regular expressions in R by Albert Kim

* Book [Handling and Processing Strings in R](http://gastonsanchez.com/Handling_and_Processing_Strings_in_R.pdf) by Gaston Sanchez
